---
title: "Final Project"
author: "Millie Meng"
date: "Spring 2026"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(car)
library(readxl)
```

The following dataset is from Joeng et al. (2017), which examined attachment styles, self-compassion, and emotional distress among 473 Korean college students. The data includes measures of anxious and avoidant attachment, self-compassion and its subscales, depression, and anxiety.

```{r load-data}
# Loading the dataset on attachment styles, self-compassion, and emotional distress
# among Korean college students (Joeng et al., 2017)
df <- read_excel("1-s2.0-S2352340917302500-mmc2.xls")

head(df)
```

Before running analyses, I converted gender and group to factor variables so R treats them as categorical. I also computed descriptive statistics for the key variables used in my analyses.

```{r data-wrangling}
# Converting gender and group to factors for analysis
df <- df %>%
  mutate(
    gender = factor(gender, levels = c(1, 2), labels = c("Male", "Female")), 
    # 1 = Male, 2 = Female; converting so R treats gender as categorical, not numeric
    group = factor(group)
  )

# Descriptive statistics for key variables
df %>%
  summarise(
    n = n(),
    M_anxious = mean(anxious_attachment),
    SD_anxious = sd(anxious_attachment),
    M_avoidant = mean(avoidant_attachment),
    SD_avoidant = sd(avoidant_attachment),
    M_selfcomp = mean(self_compassion),
    SD_selfcomp = sd(self_compassion),
    M_depression = mean(depression),
    SD_depression = sd(depression),
    M_anxiety = mean(anxiety),
    SD_anxiety = sd(anxiety)
  )
```

## Analysis 1: Regression

To examine whether attachment styles predict depression, I ran a multiple linear regression with anxious attachment and avoidant attachment as predictors and depression as the outcome variable.

```{r regression}
# Research question: Do anxious and avoidant attachment predict depression?

# Fit the regression model
lm_depression <- lm(depression ~ anxious_attachment + avoidant_attachment, data = df)
summary(lm_depression)

# Both predictors are significant (p < .001), R² = .306
# meaning attachment styles explain ~30% of variance in depression
```

I checked three assumptions of the regression model: normality of residuals (Shapiro-Wilk test), homoscedasticity (non-constant variance test), and linearity (residual plots).

```{r regression-assumptions}
# Checking assumptions of the regression model

# Assumption 1: Normality of residuals
res <- residuals(lm_depression)
shapiro.test(res)
# W = 0.969, p < .001 - normality violated, but sample size (n=473) is large enough to proceed

# Assumption 2: Homoscedasticity
ncvTest(lm_depression)
# p = .035 - mild heteroscedasticity detected, interpret results with caution

# Assumption 3: Linearity
residualPlots(lm_depression)
# Both predictors show slight non-linearity (p < .05)
```

The following plot visualizes the relationship between anxious attachment and depression with a linear regression line.

```{r regression-plot}
# Scatterplot of anxious attachment predicting depression with regression line
ggplot(df, aes(x = anxious_attachment, y = depression)) +
  geom_point(alpha = 0.3, color = "steelblue") +
  geom_smooth(method = "lm", se = TRUE, color = "navy") +
  labs(
    title = "Anxious Attachment Predicting Depression",
    x = "Anxious Attachment",
    y = "Depression"
  ) +
  theme_minimal()
```

## Analysis 2: Two-way ANOVA

To examine whether gender and university group affect anxiety levels, I ran a two-way ANOVA with gender and group as independent variables and anxiety as the outcome variable. I first computed descriptive statistics for each combination of gender and group.

```{r anova}
# Research question: Do gender and group (university) affect anxiety levels?

# Descriptive statistics by group
df %>%
  group_by(gender, group) %>%
  summarise(
    n = n(),
    M_anxiety = mean(anxiety),
    SD_anxiety = sd(anxiety),
    .groups = "drop"
  )

# Two-way ANOVA
aov_model <- aov(anxiety ~ gender * group, data = df)
summary(aov_model)

# None of the effects are significant (all p > .05)
# No post-hoc tests needed given non-significant interaction
```

I checked the assumptions of normality (Shapiro-Wilk test on residuals) and homogeneity of variance (Levene's test) for the ANOVA model.

```{r anova-assumptions}
# Normality of residuals
aov_res <- residuals(aov_model)
shapiro.test(aov_res) 
# p = .001 - mild violation, but unlikely to affect results given large sample size

# Homoscedasticity (Levene's test)
leveneTest(anxiety ~ gender * group, data = df) 
# p = .731 - assumption met; variances are equal across groups
```

The following interaction plot shows mean anxiety levels across university groups, separated by gender, with error bars representing standard error.

```{r anova-plot}
# Interaction plot: gender x group on anxiety
df %>%
  group_by(gender, group) %>%
  summarise(
    M = mean(anxiety),
    SE = sd(anxiety) / sqrt(n()),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = group, y = M, color = gender, group = gender)) +
  geom_point(size = 3) +
  geom_line() +
  geom_errorbar(aes(ymin = M - SE, ymax = M + SE), width = 0.1) +
  labs(
    title = "Anxiety by Gender and University Group",
    x = "University Group",
    y = "Mean Anxiety",
    color = "Gender"
  ) +
  theme_minimal()
```

## Analysis 3: Independent Samples T-test

To examine whether males and females differ in self-compassion, I ran an independent samples t-test. I first checked homogeneity of variance using Levene's test.

```{r ttest}
# Research question: Do males and females differ in self-compassion?

# Descriptive statistics by gender
df %>%
  group_by(gender) %>%
  summarise(
    n = n(),
    M = mean(self_compassion),
    SD = sd(self_compassion)
  )

# Check assumption: homogeneity of variance
leveneTest(self_compassion ~ gender, data = df)
# p = .060 - assumption met; proceeding with var.equal = TRUE

# Independent samples t-test
t.test(self_compassion ~ gender, data = df, var.equal = TRUE)
# t = -0.055, p = .956 - no significant difference between males and females
# in self-compassion (Males: M = 3.14, Females: M = 3.14)
```

The following violin plot shows the distribution of self-compassion scores for males and females.

```{r ttest-plot}
# Violin plot comparing self-compassion by gender
ggplot(df, aes(x = gender, y = self_compassion, fill = gender)) +
  geom_violin(alpha = 0.5) +
  geom_boxplot(width = 0.1, alpha = 0.8) +
  labs(
    title = "Self-Compassion by Gender",
    x = "Gender",
    y = "Self-Compassion"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```
